---
title: The Civil Society Organizations effect: A mixed-Methods analysis of bottom-up approaches in Brazilian public policy
runninghead: Galdino_et al._.
author:
- name: Manoel Galdino*
  num: 1,2
- name: Juliana Sakai
  num: 2
- name: Bianca Vaz Mondo
  num: 3
- name: Natália Paiva
  num: 4
address:
- num: 1
  org: Department of Political Science, Universidade of São Paulo, Brazil
- num: 2
  org: Transparência Brasil, Brazil
- num: 3
  org: Herthie School, Germany
- num: 4
  org: Alandar Consulting, Brazil
corrauth: "Manoel Galdino, Avenida Luciano Gualberto, 315.Sala 2081. CEP 05508-090. São Paulo - SP, Brazil."
email: mgaldino@usp.br
abstract: "This paper examines the effect of bottom-up accountability on public service delivery. We differentiate between information-driven interventions and the mobilization and monitoring efforts of organized Civil Society Organizations (CSOs), and argue that the latter type of interventions can can drive significant policy change. The study evaluates the effectiveness of the Obra Transparente project by Brazilian NGO Transparência Brasil, engaging 21 local CSOs in South and Southeast Brazil.  Our findings emphasize the importance of sustained, coordinated efforts by socially embedded CSOS. These efforts, involving direct engagement with municipal officials and ensuring that their complains cannot be ignored lead to more substantial outcomes when compared with citizens information-driven"
keywords: Class file; \LaTeX; SMMR; Rmarkdown;
classoption:
  - Royal
  - times
bibliography: OT_paper.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
---

# The Article Header Information

YAML header:

```
output:
  rticles::sim_article:
    keep_tex: TRUE
```

Configure the YAML header including the following elements:

* `title`: Title

* `runninghead`: Author last names, use _et al._ if there are three or more authors.

* `author`: List of author(s) containing `name` and `num`

* `corrauth`: Corresponding author's name and address.

* `email`: Correspondence email

* `abstract`: Limited to 200 words

* `keywords`: Keywords for the article

* `bibliography`: BibTeX `.bib` file

* `bibliographystyle`: sageh or sagev

* `classoption`: options of the `sagej` class

## Remarks

2. `bibliographystyle`

3. `classoption`

4. Keywords are separated by commas.

# The Body of the Article

## Mathematics

Use mathematics in Rmarkdown as usual.

## Figures and Tables

Figures are supported from R code:

```{r plot-ref, fig.cap = "Fancy Caption\\label{fig:plot}", out.width='100%'}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
```

...and can be referenced (Figure \ref{fig:plot}) by including the `\\label{}` tag in the `fig.cap` attribute of the R chunk: `fig.cap = "Fancy Caption\\label{fig:plot}"`. It is a quirky hack at the moment, see [here](https://github.com/yihui/knitr/issues/323).

Analogously, use Rmarkdown to produce tables as usual:

```{r, results = "asis"}
if (!require("xtable")) install.packages("xtable")
xt <- xtable(head(cars), caption = "A table", label = "tab:table")
print(xt, comment = FALSE)
```

Referenced via \ref{tab:table}. @article{gelmanWeaklyInformativeDefault2008,
  title = {A Weakly Informative Default Prior Distribution for Logistic and Other Regression Models},
  author = {Gelman, Andrew and Jakulin, Aleks and Pittau, Maria Grazia and Su, Yu-Sung},
  date = {2008}
}
It is also possible to set the `YAML` option `longtable: true` and use markdown tables (or the `knitr::kable` function): `knitr::kable(head(cars))` produces the same table as the `xtable` example presented before.

## Cross-referencing

The use of the Rmarkdown equivalent of the \LaTeX cross-reference system
for figures, tables, equations, etc., is encouraged (using `[@<name>]`, equivalent of `\ref{<name>}` and `\label{<name>}`). That works well for citations in Rmarkdown, not so well for figures and tables. In that case, it is possible to revert to standard \LaTeX syntax.

## Double Spacing

If you need to double space your document for submission please use the `doublespace` option in the header.

# Bibliography

Link a `.bib` document via the YAML header, and bibliography will be printed at the very end (as usual). The default bibliography style is provided by Wiley as in `WileyNJD-AMA.bst`, do not delete that file.

Use the Rmarkdown equivalent of the \LaTeX citation system using `[@<name>]`. Example: [@Taylor1937], [@Knupp1999; @Kamm2000].

To include all citation from the `.bib` file, add `\nocite{*}` before the end of the document.

# Further information

All \LaTeX enviroments supported by the main template are supported here as well; see the `.tex` sample file [here](http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1097-0258/homepage/la_tex_class_file.htm) for more details and example.
